Concurrent Programming in Harmony

### 2021-08-08 book/paper.tex
Latest

The Needham-Schroeder protocol~\cite{NS78} is a security protocol in which two parties
authenticate one another by exchanging large and recently created random numbers
called \emph{nonces} that nobody else should be able to read.
The version we describe here uses \emph{public key cryptography}~\cite{pkc}: with public key
cryptography it is possible to create a message for a particular destination that
only that destination can read.  We denote with $\langle m \rangle_p$ a message
$m$ encrypted for $p$ so that only $p$ can decrypt the message and see that it contains $m$..

The original protocol goes something like this.  Suppose Alice wants to communicate
with Bob.  The three steps are as follows:

@@

When Bob receives $N_B$ in the last step, it decides that it is Alice that is
trying to communicate at this time.  Since it created $N_B$ recently and sent it
encrypted to Alice, Bob does not have to worry that the message was an old message
that was replayed.  Also, if Alice is honest, only Alice can have seen the message.

\autoref{fig:ns} shows the protocol implemented in Harmony.
A message $\langle m \rangle_p$ is encoded in Harmony as a dictionary
$\{ \mathtt{.dst}: p, \mathtt{.contents}: m \}$.

Unfortunately, the protocol turns out to be incorrect, but it took 17 years before
somebody noticed~\cite{Lowe95}.  The bug is easily found using model
checking~\cite{Lowe96}.

>>>>

The Needham-Schroeder protocol~\cite{NS78} is a security protocol in which two parties
authenticate one another by exchanging large and recently created random numbers
called \emph{nonces} that nobody else should be able to read.
The version we describe here uses \emph{public key cryptography}~\cite{DH76}: with public key
cryptography it is possible to create a message for a particular destination that
only that destination can read.  We denote with $\langle m \rangle_p$ a message
$m$ encrypted for $p$ so that only $p$ can decrypt the message and see that it contains $m$.

Suppose Alice wants to communicate with Bob.  The three critical steps in the
Needham-Schroeder protocol are as follows:

@@

When Bob receives $N_B$ in the last step, it decides that it is Alice that is
trying to communicate at this time.  Since Bob created $N_B$ recently and sent it
encrypted to Alice, Bob does not have to worry that the message was an old message
that was replayed.  Also, if Alice is honest, only Alice can have seen the message
containing $N_B$.

\autoref{fig:ns} shows the protocol implemented in Harmony.
A message $\langle m \rangle_p$ is encoded in Harmony as a dictionary
$\{ \mathtt{.dst}: p, \mathtt{.contents}: m \}$.  The code for Alice and Bob
simply follows the steps listed above.

Unfortunately, the protocol turns out to be incorrect, but it took 17 years before
somebody noticed~\cite{Lowe95}.  Model checking can be used to find the bug~\cite{Lowe96}.
To demonstate the bug, we need a third party, which we will call Carter.
We want to make sure that Carter cannot impersonate Alice or Bob.
However, it is possible that Alice tries to set up an authenticated connection to Carter
using the Needham-Schroeder protocol.  That in itself is not a problem.

The code in \autoref{fig:ns} has Alice try to set up a connection to either Bob
or Carter.  Bob only accepts connections with Alice.  Carter, when receiving a message
that it can decrypt, will try to find an attack by sending every possible message to
every possible destination.  In particular, it keeps track of every nonce that it has
seen and will try to construct messages with them to send to Alice and Bob.
If, at some point, Carter finds the nonce that Bob generates for Alice, then Carter
can impersonate Alice by sending a type~3 message containing that nonce to Bob.

Running the code in \autoref{fig:ns} quickly finds a viable attack.  The attack goes
like this:

\begin{enumerate}
\item Alice creates a new nonce $N_A$ and sends $\langle 1, A, N_A \rangle_\mathtt{Carter}$
to Carter;
\item Upon receipt, Carter sends $\langle 1, A, N_A \rangle_\mathtt{Bob}$ to Bob;
\item Upon receipt, Bob creates a new nonce $N_B$ and sends
$\langle 2, N_A, N_B \rangle_\mathtt{Alice}$ to Alice;
\item Alice thinks the message came from Carter (because it contains $N_A$, which
Alice sent to Carter) and
sends $\langle 3, N_B \rangle_\mathtt{Carter}$ to Carter.
\item Carter learns $N_B$, which it was not supposed to.
\end{enumerate}

Carter can use $N_B$ to send $\langle 3, N_B \rangle_\mathtt{Bob}$ to Bob.
Bob now thinks it is talking to Alice, even though Alice never tried to
communicate with Bob.
